My title

### 2024-10-18 paper/paper.qmd
finished the graphs for the paper

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Mandy He
  - Wendy Yuan
thanks: "Code and data are available at: https://github.com/MandyHe7/US-Election.git."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(palmerpenguins)
```

@@

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

@@

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

@@

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

2. Population, Frame, and Sample

Population: YouGov’s target population consists of U.S. likely voters, defined by their eligibility and likelihood to vote, assessed through past voting behavior, voter registration, and expressed intent.

Frame: The sampling frame is YouGov’s online panel, which includes millions of U.S. residents recruited through various methods. The panel is stratified and weighted to represent key demographics such as age, gender, race, education, and region.

Sample: YouGov uses non-probability sampling from its panel, adjusting for biases using sophisticated weighting to reflect the U.S. electorate's composition.

3. Sample Recruitment

YouGov recruits panelists through:
Targeted online ads: Ads across websites and social media attract diverse participants.
Affiliate marketing: Collaborations with other websites help reach a broad audience.
Incentives: Respondents earn points for surveys, increasing participation and retention.
However, because respondents voluntarily join the panel, there is potential for self-selection bias, as certain demographics may be more likely to participate.

>>>>

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - Mandy He
  - Wendy Yuan
thanks: "Code and data are available at: https://github.com/MandyHe7/US-Election.git."
date: November 5, 2024
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

library(tidyverse)
library(dplyr)
library(readr)
library(ggplot2)
library(janitor)
library(lubridate)
library(broom)
library(modelsummary)
library(rstanarm)
library(splines)

polling_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
```

@@

@@

@@

```{r}
#### Plot data ####
base_plot <- ggplot(polling_data, aes(x = end_date, y = pct, color = answer)) +
  geom_point() +  # Add points to the plot
  theme_classic() +
  labs(y = "Percent", x = "Date", color = "Candidate") +  # Add a label for the legend
  scale_color_manual(values = c("Biden" = "blue", "Harris" = "orange"))  # Customize colors

# Plots poll estimates and overall smoothing
base_plot +
  geom_point() +
  geom_smooth()

```

```{r}
# Facet by pollster
base_plot +
  geom_point() +
  geom_smooth() +
  facet_wrap(vars(pollster))
```

```{r}
# Facet by state
base_plot +
  geom_point() +
  geom_smooth() +
  facet_wrap(vars(state))
```

@@

```{r}
#### Bayesian modeling ####
# Change 'pollster' and 'state' to factor variables
polling_data <- polling_data |>
  mutate(
    pollster = factor(pollster),
    state = factor(state)
  )

# Check if 'num_DEM' and 'sample_size' exist; if not, create them (adjust based on actual column names)
# Assuming 'pct_DEM' is the percentage of Democrat votes (adjust if necessary)
if (!("num_DEM" %in% colnames(polling_data))) {
  polling_data <- polling_data |> mutate(num_DEM = round(pct_DEM * sample_size / 100))
}

# Model 1
model_formula_1 <- cbind(num_DEM, sample_size - num_DEM) ~ (1 | pollster)

# Specify priors
priors <- normal(0, 2.5, autoscale = TRUE)

# Fit the models
bayesian_model_1 <- stan_glmer(
  formula = model_formula_1,
  data = polling_data,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 1,
  adapt_delta = 0.99  # Increase adapt_delta to avoid divergent transitions
)

# Posterior predictive checks
pp_check(bayesian_model_1)

#### Bayesian models and splines ####
# Change date to be number of days since she declared - it's a counter not a date
polling_data <- polling_data |>
  mutate(
    end_date_num = as.numeric(end_date - min(end_date))
  )

# Fit Bayesian model with spline and pollster as fixed effect
spline_model <- stan_glm(
  pct ~ ns(end_date, df = 5) + pollster,  # Change df for the number of "bits"
  data = polling_data,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

# Posterior predictive checks
pp_check(spline_model)

# Predict and plot
new_data <- data.frame(
  end_date = seq(
    min(polling_data$end_date),
    max(polling_data$end_date),
    length.out = 100
  ),
  pollster = factor("YouGov", levels = levels(polling_data$pollster))
)

# Predict posterior draws
posterior_preds <- posterior_predict(spline_model, newdata = new_data)

# Summarize predictions
pred_summary <- new_data |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
  )

# Plot the spline fit
ggplot(polling_data, aes(x = end_date, y = pct, color = pollster)) +
  geom_point() +
  geom_line(
    data = pred_summary,
    aes(x = end_date, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE
  ) +
  geom_ribbon(
    data = pred_summary,
    aes(x = end_date, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit"
  ) +
  theme_minimal()
```

```{r}
#### Bayesian modeling ####
# Change 'pollster' and 'state' to factor variables
polling_data_2 <- polling_data |>
  mutate(
    pollster = factor(pollster),
    state = factor(state)  # State as a factor variable
  )

# Check if 'num_DEM' and 'sample_size' exist; if not, create them (adjust based on actual column names)
# Assuming 'pct_DEM' is the percentage of Democrat votes (adjust if necessary)
if (!("num_DEM" %in% colnames(polling_data))) {
  polling_data_2 <- polling_data_2 |> mutate(num_DEM = round(pct_DEM * sample_size / 100))
}

# Model 1: Random effect for state instead of pollster
model_formula_2 <- cbind(num_DEM, sample_size - num_DEM) ~ (1 | state)

# Specify priors
priors <- normal(0, 2.5, autoscale = TRUE)

# Fit the model with random effect for state
bayesian_model_2 <- stan_glmer(
  formula = model_formula_1,
  data = polling_data,
  family = binomial(link = "logit"),
  prior = priors,
  prior_intercept = priors,
  seed = 123,
  cores = 1,
  adapt_delta = 0.99  # Increase adapt_delta to avoid divergent transitions
)

# Posterior predictive checks
pp_check(bayesian_model_2)

#### Bayesian models and splines ####
# Change date to be number of days since she declared - it's a counter not a date
polling_data_2 <- polling_data_2 |>
  mutate(
    end_date_num = as.numeric(end_date - min(end_date))
  )

# Fit Bayesian model with spline and random effect for state
spline_model_2 <- stan_glm(
  pct ~ ns(end_date, df = 5) + state,  # State as random effect
  data = polling_data_2,
  family = gaussian(),
  prior = normal(0, 5),
  prior_intercept = normal(50, 10),
  seed = 1234,
  iter = 2000,
  chains = 4,
  refresh = 0
)

# Posterior predictive checks
pp_check(spline_model_2)

# Predict and plot
new_data_2 <- data.frame(
  end_date = seq(
    min(polling_data$end_date),
    max(polling_data$end_date),
    length.out = 100
  ),
  state = factor("California", levels = levels(polling_data$state))  # Example state
)

# Summarize predictions
pred_summary_2 <- new_data_2 |>
  mutate(
    pred_mean = colMeans(posterior_preds),
    pred_lower = apply(posterior_preds, 2, quantile, probs = 0.025),
    pred_upper = apply(posterior_preds, 2, quantile, probs = 0.975)
  )

# Plot the spline fit
ggplot(polling_data_2, aes(x = end_date, y = pct, color = state)) +
  geom_point() +
  geom_line(
    data = pred_summary_2,
    aes(x = end_date, y = pred_mean),
    color = "blue",
    inherit.aes = FALSE
  ) +
  geom_ribbon(
    data = pred_summary_2,
    aes(x = end_date, ymin = pred_lower, ymax = pred_upper),
    alpha = 0.2,
    inherit.aes = FALSE
  ) +
  labs(
    x = "Days since earliest poll",
    y = "Percentage",
    title = "Poll Percentage over Time with Spline Fit"
  ) +
  theme_minimal()
```

@@

2. Population, Frame, and Sample
Population: YouGov’s target population consists of U.S. likely voters, defined by their eligibility and likelihood to vote, assessed through past voting behavior, voter registration, and expressed intent.
Frame: The sampling frame is YouGov’s online panel, which includes millions of U.S. residents recruited through various methods. The panel is stratified and weighted to represent key demographics such as age, gender, race, education, and region.
Sample: YouGov uses non-probability sampling from its panel, adjusting for biases using sophisticated weighting to reflect the U.S. electorate's composition.

3. Sample Recruitment
YouGov recruits panelists through:
Targeted online ads: Ads across websites and social media attract diverse participants.
Affiliate marketing: Collaborations with other websites help reach a broad audience.
Incentives: Respondents earn points for surveys, increasing participation and retention.
However, because respondents voluntarily join the panel, there is potential for self-selection bias, as certain demographics may be more likely to participate.
